Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control: A Secondary Analysis of SPRINT

### 2022-01-07 doc/manuscript.Rmd
bcj intro edits

both the The Systolic Blood Pressure Intervention Trial (SPRINT) and the Strategy of Blood Pressure Intervention in the Elderly Hypertensive Patients (STEP) trial have shown the benefits of intensive blood pressure control on cardiovascular morbidity and mortality [@sprint_research_group_randomized_2015; @zhang_trial_2021]. However, both trials were stopped early after a median follow-up of 3 years.

__Reminder: bring in the recent BPTTC meta-analyses into the introduction__[@rahimi_age_stratified_2021].

Here we examined the longer term effect of randomization to intensive treatment on through linkage to the National Death Index (NDI).

@@

__Trial Design__: The trial design and methods have been published previously [@ambrosius_design_2014]. Briefly, we conducted a multicenter randomized clinical trial that compared two strategies for managing systolic BP (SBP) in older adults with hypertension who were at increased risk of cardiovascular disease (CVD). Participants were aged 50 years or older and had an SBP between 130 and 180 mm Hg at the screening visit, depending on the number of anti-hypertensive agents prescribed. Participants were considered to have an increased cardiovascular risk if they had clinical or subclinical cardiovascular disease, chronic kidney disease (defined by an estimated glomerular filtration rate of <60 mL/min/1.73 m^2^), or a Framingham Risk Score of 15% or greater or if they were aged 75 years or older. Individuals residing in a nursing home, persons with a diagnosis of dementia (based on medical record review), and those treated with medications primarily used for dementia therapy were excluded, as were persons with prevalent diabetes mellitus, history of stroke, proteinuria > 1 gram per day, or polycystic kidney disease. Individuals at 102 sites in the United States and Puerto Rico were randomized (1:1) to a SBP goal of less than 120 mm Hg (intensive treatment group, n = 4678) or a goal of less than 140 mm Hg (standard treatment group, n = 4683), using random permuted blocks with the randomization stratified by clinic site. The algorithms and formulary for the trial are listed in the published study protocol ( __citations needed__? ). Trial enrollment began in November 2010 and ended in March 2013, with active follow-up through July 1, 2016. The study was approved by the institutional review board at each participating site, and each participant provided written informed consent. The study is registered at ClinicalTrials.gov (NCT01206062).

>>>>

Blood pressure (BP) is the leading modifiable risk factor for cardiovascular disease (CVD). A meta analysis of pooled clinical trial data has shown that pharmacological BP reduction reduces the risk of major cardiovascular events, with larger absolute risk reduction for older adults [@rahimi_age_stratified_2021]. The Systolic Blood Pressure Intervention Trial (SPRINT) and the Strategy of Blood Pressure Intervention in the Elderly Hypertensive Patients (STEP) each showed that intensive systolic BP (SBP) control, defined by a SBP < 120 mm Hg in SPRINT and SBP < 130 mm Hg in STEP, reduced the risk for cardiovascular morbidity and mortality [@sprint_research_group_randomized_2015; @zhang_trial_2021]. However, as both trials were stopped early after a median follow-up of 3 years, the longer term effect of randomization to intensive SBP control on CVD risk and SBP levels following the trial period are unclear.

The objective of the current study was to estimate the longer term effect of randomization to intensive SBP control on CVD risk for SPRINT participants. A secondary objective was to describe SBP levels over time following the trial period. To accomplish these objectives, we linked SPRINT participant data to the National Death Index (NDI) and pulled longitudinal SBP data for participants with available BP values in the electronic health record (EHR).

@@

__Trial Design__: The design and methods of SPRINT have been published previously [@ambrosius_design_2014]. Briefly, we conducted a multicenter randomized clinical trial that compared two strategies for managing SBP in older adults with hypertension who were at increased risk of CVD. Participants were aged 50 years or older and had an SBP between 130 and 180 mm Hg at the screening visit, depending on the number of anti-hypertensive agents prescribed. Participants were considered to have an increased cardiovascular risk if they had clinical or subclinical cardiovascular disease, chronic kidney disease (defined by an estimated glomerular filtration rate of <60 mL/min/1.73 m^2^), or a Framingham Risk Score of 15% or greater or if they were aged 75 years or older. Individuals residing in a nursing home, persons with a diagnosis of dementia (based on medical record review), and those treated with medications primarily used for dementia therapy were excluded, as were persons with prevalent diabetes mellitus, history of stroke, proteinuria > 1 gram per day, or polycystic kidney disease. Individuals at 102 sites in the United States and Puerto Rico were randomized (1:1) to a SBP goal of less than 120 mm Hg (intensive treatment group, n = 4678) or a goal of less than 140 mm Hg (standard treatment group, n = 4683), using random permuted blocks with the randomization stratified by clinic site. The algorithms and formulary for the trial are listed in the published study protocol ( __citations needed__? ). Trial enrollment began in November 2010 and ended in March 2013, with active follow-up through July 1, 2016. The study was approved by the institutional review board at each participating site, and each participant provided written informed consent. The study is registered at ClinicalTrials.gov (NCT01206062).
